---
title: "R Notebook"
output: html_notebook
---

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). I have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

N.B. Question to be answered via email to Celine: can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.

N.B. There are looots of files to be dealt with. Probably too many for your computer. This is a challenge for you. Some (complementary) possible strategies: You can select a subset of files only (and you have to justify your choice). You can learn how to use the apply() or map() functions. You can coordinate with classmates.

1. In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)
- Extract "complex" descriptors: recurrence quantification analysis

```{r}
# Set working directory
setwd("~/OneDrive - Aarhus universitet/AU-Cognitive Science/3rd Semester/Experimental Methods 3/Exercise/Assignments/Assignment_2")

# Loading libraries
library(raster); library(lsr); library(crqa); library(ggplot2)

# Load in the first datafile
data1=read.delim("Pitch/Study1D0S102T1_f0.txt")

# Load demodata 
demo_data = read.delim("DemoData.txt")

# To extract mean
mean(data1$f0) 

# SD
sd(data1$f0)

# Range
range(data1$f0)

# median
median(data1$f0)

# IQR
IQR(data1$f0)

# Mean absoluted deviation
aad(data1$f0)

# Coefficient of variation
sd(data1$f0)/mean(data1$f0)*100

# Recurrence Quantificantion Analysis
# To estimate paramters (Dimension, Radius and delay parameter)
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100, radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2, minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE, fnnpercent = 10,  typeami = "mindip") 

ans= optimizeParam(data1$f0, data1$f0, par, min.rec= 3.5, max.rec=4.5)

results=crqa(data1$f0, data1$f0, delay=ans$delay, embed=ans$emddim, radius=ans$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)

# To view the plot
RP=results$RP 
RP = matrix(as.numeric(RP), nrow = ncol(RP))  
cols = c("white","blue4") 
image(RP, xlab = "", ylab = "", col = cols)

# Explore lags of coordination 
Profile=drpdfromts(data1$f0, data1$f0, datatype='continuous', ws=50,radius=ans$radius) 
timecourse = round( seq(-5000,5000,100)/1000, digit = 1) 
maxlag = Profile$maxlag/1000 
profile = Profile$profile*100 
Prof=data.frame(profile) 
ggplot(Prof, aes(timecourse,profile))+geom_line()+geom_vline(xintercept = timecourse[maxlag], colour='red')

```

```{r}
# To create empty variables
filename= NULL
mean = NULL
SD = NULL
range = NULL
median = NULL
IQR = NULL
MAD = NULL
COV = NULL
DET = NULL
maxL = NULL
RR = NULL
L = NULL
ENTR = NULL
rENTR = NULL

n = 1

# To create a filelist
files=list.files(path = "Pitch", pattern = ".txt")

# To set working directory to file location
setwd("~/OneDrive - Aarhus universitet/AU-Cognitive Science/3rd Semester/Experimental Methods 3/Exercise/Assignments/Assignment_2/Pitch")

# To loop through each file and compute descriptors
for (file in files){
  d=read.delim(file)
  filename[n] = files[n]
  mean[n]=mean(d$f0)
  SD[n]=sd(d$f0)
  range[n]=range(d$f0)
  median[n] = median(d$f0)
  IQR[n] = IQR(d$f0)
  MAD[n] = aad(d$f0)
  COV[n] = sd(d$f0)/mean(d$f0)*100
  
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,
  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,
  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,
  fnnpercent = 10,  typeami = "mindip") 

  ans= try(optimizeParam(d$f0, d$f0, par, min.rec= 3.5,
                     max.rec=4.5))
  if (length(ans) > 1) {
  
  results=crqa(d$f0, d$f0, delay=ans$delay, embed=ans$emddim,
               radius=ans$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
  maxL[n] = results$maxL
  DET[n] = results$DET
  RR[n] = results$RR
  L[n] = results$L
  ENTR[n] = results$ENTR
  rENTR[n] = results$rENTR
  } 
  
  else {
    DET[n] = NA
    maxL[n] = NA
    RR[n] = NA
    L[n] = NA
    ENTR[n] = NA
    rENTR[n] = NA
  }
  
  n=n+1
}


# To create dataframe with pitch descriptors
Pitch_data=data.frame(filename, mean, SD, range, median, IQR, MAD, COV, DET, maxL, RR, L, ENTR, rENTR)
Pitch_data = read.csv("Pitch_data.csv")
----


# To create a variable according to the study
for (file in 1:nrow(Pitch_data)){
  if (grepl("Study1", Pitch_data$filename[file])){
    Pitch_data$Study[file] = 1}
  if (grepl("Study2", Pitch_data$filename[file])){
    Pitch_data$Study[file] = 2}
  if (grepl("Study4", Pitch_data$filename[file])){
    Pitch_data$Study[file] = 4}
}

# To create a variable according to diagnosis
for (file in 1:nrow(Pitch_data)){
  if (grepl("D0", Pitch_data$filename[file])){
    Pitch_data$Diagnosis[file] = "Control"}
  if (grepl("D1", Pitch_data$filename[file])){
    Pitch_data$Diagnosis[file] = "Schizophrenia"}
}

# To create a variable according to trial
Pitch_data$Trial=regmatches(Pitch_data$filename, regexpr("[T].*[0-9]", Pitch_data$filename))

Pitch_data$Trial = gsub("[T, _, f]", "", Pitch_data$Trial)

Pitch_data$Trial=substr(Pitch_data$Trial, 1, nchar(Pitch_data$Trial)-1)

# To create a variable according to ID
Pitch_data$Subject = substring(Pitch_data$filename, 2)

Pitch_data$Subject=regmatches(Pitch_data$Subject, regexpr("[S].*[0-9].[T]", Pitch_data$Subject))

Pitch_data$Subject = gsub("[A-Z]", "", Pitch_data$Subject)

# To merge the pitch data together with the demo data
Pitch_data$Diagnosis=as.factor(Pitch_data$Diagnosis)
Pitch_data$Subject=as.integer(Pitch_data$Subject)

Pitch_data = read.csv("Pitch_data.csv")

new_data = base::merge(Pitch_data, demo_data, by = c("Subject", "Diagnosis"))

# To write to a csv file
write.csv(Pitch_data, "Pitch_data.csv")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

