---
title: "R Notebook"
output: html_notebook
---

## Assignment 3 - Diagnosing schizophrenia from voice
In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

### Question 1: Can you diagnose schizophrenia from pitch range only? If so, how well?
Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
```{r}

setwd("~/OneDrive - Aarhus universitet/AU-Cognitive Science/3rd Semester/Experimental Methods 3/Exercise/Assignments/Assignment_2")
data=read.csv("Schizophrenia_data.csv")

library(pastecs);library(lme4);library(lmerTest);library(ggplot2);library(caret);library(pROC);library(tidyverse);library(caret);library(cvms); library(groupdata2) 

#Reading libraries
library(lmerTest)
library(caret)
library(pROC)

#Reading the data
data$Subject=as.factor(data$Subject)

#Making a generalized linear model
model1=glmer(Diagnosis~range + (1+Trial|Subject), data, family="binomial")
summary(model1)

#Creating confusion matrix
data$PredictionsPerc=predict(model1)
data$Predictions[data$PredictionsPerc>0]="Schizophrenia"
data$Predictions[data$PredictionsPerc<=0]="Control"
c=confusionMatrix(data = data$Predictions, reference = data$Diagnosis, positive = "Schizophrenia") 


#Making a reciever operator curve
rocCurve=roc(response=data$Diagnosis, predictor=data$PredictionsPerc)
auc(rocCurve)
ci(rocCurve)
plot(rocCurve, legacy.axes=TRUE)

#Cross-validating model

# Adding fold_value
fold_function = function(data, col) {

  data$folds_ID = as.numeric(interaction(data[,col]))
  
  return(data[order(data$folds_ID),])
}

Data_Schizo = subset(data, data$Diagnosis == "Schizophrenia")
Data_Control = subset(data, data$Diagnosis == "Control")

Data_Schizo=fold_function(Data_Schizo, "Subject")
Data_Control= fold_function(Data_Control, "Subject")

# Creating folds
folds_S = createFolds(unique(Data_Schizo$folds_ID), 5)
folds_C= createFolds(unique(Data_Control$folds_ID), 5)

# Combining data
fold_list = c()
fold_list$Fold1 = c(folds_S$Fold1, folds_C$Fold1)
fold_list$Fold2 = c(folds_S$Fold2, folds_C$Fold2)
fold_list$Fold3 = c(folds_S$Fold3, folds_C$Fold3)
fold_list$Fold4 = c(folds_S$Fold4, folds_C$Fold4)
fold_list$Fold5 = c(folds_S$Fold5, folds_C$Fold5)

NewData = rbind(Data_Schizo, Data_Control)

#Creating a loop

Accuracy=NULL
Specificity=NULL
Sensitivity=NULL
PPV=NULL
NPV=NULL
AUC=NULL

n=1

for (fold in fold_list){
  #Creating a second training dataset 
  train=subset(NewData,! (folds_ID %in% fold))
  
  #Creaing a second test dataset
  test=subset(NewData, (folds_ID %in% fold))
  
  #Creating a model - train
  TrainModel=glmer(Diagnosis~range + (1+Trial|Subject), train, family="binomial")
  
  test$PredictionsPerc=predict(TrainModel, test, allow.new.levels=TRUE)
  test$Predictions[test$PredictionsPerc>0]="Schizophrenia"
  test$Predictions[test$PredictionsPerc<=0]="Control"
  
  cm=confusionMatrix(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 
  
  Accuracy[n]=cm$overall["Accuracy"]
  
  test$Predictions=as.factor(test$Predictions)
  
  Specificity[n]= specificity(data = test$Predictions, reference = test$Diagnosis, negative = "Control") 
  Sensitivity[n]= sensitivity(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia")
  PPV[n]= posPredValue(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia")
  NPV[n]= negPredValue(data = test$Predictions, reference = test$Diagnosis, negative = "Control") 
  
  rocCurve=roc(response=test$Diagnosis, predictor=test$PredictionsPerc)
  AUC[n]=auc(rocCurve)
    
  n=n+1
 
}

TestData=data.frame(Accuracy, Specificity, Sensitivity, PPV, NPV, AUC)

```

### Question 2
Which single predictor is the best predictor of diagnosis?
```{r}
# Creating a function to find out which single acoustic predictor is the best predictor
CVfunction = function(data, model) {

# Creating empty variables
Accuracy = NULL
Specificity = NULL
Sensitivity = NULL
PPV = NULL
NPV = NULL
AUC = NULL

n = 1

# Creating a loop
for (fold in fold_list){
  
  # Creating training data set
  train = subset(data,! (folds_ID %in% fold))  
  
  # Creating a test data set
  test = subset(data, (folds_ID %in% fold)) 
  
  # Training model
  m1=glmer(model, train, family = "binomial")
  
  # Evaluating model
  test$PredictionsPerc=predict(m1, test, allow.new.levels = TRUE) 
  test$Predictions[test$PredictionsPerc>0]="Schizophrenia" 
  test$Predictions[test$PredictionsPerc<=0]="Control"
  
  cm=confusionMatrix(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 
  
  Accuracy[n] = cm$overall["Accuracy"]
  
  test$Predictions = as.factor(test$Predictions)

  Sensitivity[n] =  sensitivity(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia" )
  
  Specificity [n] = specificity(data = test$Predictions, reference = test$Diagnosis, negative = "Control" ) 

  PPV[n] = posPredValue(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 

  NPV[n] = negPredValue(data = test$Predictions, reference = test$Diagnosis, negative = "Control")  
  
  rocCurve = roc(response = test$Diagnosis,   predictor = test$PredictionsPerc) 
  AUC[n]=auc(rocCurve) 
    
  n=n+1  
}
results = data.frame(Accuracy, Sensitivity, Specificity, PPV, NPV, AUC)
return(results)
}

MeanResults = CVfunction(NewData, Diagnosis ~ mean + (1+Trial|Subject))
SDresults = CVfunction(NewData, Diagnosis ~ SD + (1+Trial|Subject))
MedianResults = CVfunction(NewData, Diagnosis ~ median + (1+Trial|Subject))
RRresults = CVfunction(NewData, Diagnosis ~ RR + (1+Trial|Subject))
ENTRresults = CVfunction(NewData, Diagnosis ~ ENTR + (1+Trial|Subject))
rENTRresults = CVfunction(NewData, Diagnosis ~ rENTR + (1+Trial|Subject))

# Obtaining the mean area under the curve
mean(SDresults$AUC)
# 0.61
mean(MeanResults$AUC) 
# 0.57
mean(MedianResults$AUC) 
# 0.57
mean(RRresults$AUC)
# 0.51
mean(ENTRresults$AUC) 
# 0.61
mean(rENTRresults$AUC) 
# 0.56

```

