---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2017"
output: html_document
---

```{r setup, include=FALSE}
# set WD
setwd("~/OneDrive - Aarhus universitet/AU-Cognitive Science/3rd Semester/Experimental Methods 3/Exercise/Assignments/Assignment_2")

#library to relevant packages
library(metafor);library(pacman);library(lmerTest)

#read data
data=read.csv("Data.csv", sep=";")
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered
###Exercise 1
1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 

######Yi=Cohen's D
######Vi=variance of Cohen's D
```{r}
#using escalc for effectsize
PitchRangeM=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = data)

# making lm
ModelM = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=PitchRangeM, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(ModelM)

(scipen=3) 
resM = rma(yi, vi, data = PitchRangeM, slab=Article)
summary(resM)
forest(resM)

SubsetM=subset(PitchRangeM[PitchRangeM$vi>0,])
SCHIpar=sum(SubsetM$SampleSizeSchizo, na.rm=T)
CONRpar=sum(SubsetM$SampleSizeContros, na.rm=T)
```
####Reporting
4 studies
166 participants with schizophrenia and 95 comparison participants.
The overall estimated difference (Cohen’s d) in pitch variability between the ASD and the comparison groups was 0.18, SE= 0.17

```{r}
#using escalc for standarderror
PitchRangeSD=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = data)

# making lm
ModelSD = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=PitchRangeSD, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(ModelSD)

resSD = rma(yi, vi, data = PitchRangeSD, slab=Article)
summary(resSD)
forest(resSD)

SubsetSD=subset(PitchRangeSD[PitchRangeSD$vi>0,])
SCHIparSD=sum(SubsetSD$SampleSizeSchizo, na.rm=T)
CONRparSD=sum(SubsetSD$SampleSizeContros, na.rm=T)
```
####Reporting
6 studies
257 participants with schizophrenia and 156 comparison participants.
The overall estimated difference (Cohen’s d) in pitch variability between the schizophrenia and the comparison groups was -0.13, SE= 0.58

###Exercise 
2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.
```{r}
#Pitch_data = read.csv("AllData.csv")
Pitch_data=read.csv("Schizophrenia_data.csv")
# Pitch_data = read.csv("dataPitch.csv")
Pitch_data = Pitch_data[,-1]
#mean
PitchMeanSchizo=mean(Pitch_data$mean[Pitch_data$Diagnosis=="Schizophrenia"])
PitchMeanControls=mean(Pitch_data$mean[Pitch_data$Diagnosis=="Control"])
PitchMeanSchizoSD=sd(Pitch_data$mean[Pitch_data$Diagnosis=="Schizophrenia"])
PitchMeanControlsSD=sd(Pitch_data$mean[Pitch_data$Diagnosis=="Control"])
#SD
PitchSDSchizo=mean(Pitch_data$SD[Pitch_data$Diagnosis=="Schizophrenia"])
PitchSDControls=mean(Pitch_data$SD[Pitch_data$Diagnosis=="Control"])
PitchSDSchizoSD=sd(Pitch_data$SD[Pitch_data$Diagnosis=="Schizophrenia"])
PitchSDControlsSD=sd(Pitch_data$SD[Pitch_data$Diagnosis=="Control"])

#Counting participants from own dataset
SampleSizeSchizo=length(unique(Pitch_data$Subject[Pitch_data$Diagnosis=="Schizophrenia"]))
SampleSizeContros=length(unique(Pitch_data$Subject[Pitch_data$Diagnosis=="Control"]))
Article="Prev. Schizophrenia assignment"
Year=2017
data11=data.frame(Article,Year,SampleSizeSchizo,SampleSizeContros,PitchMeanControls,PitchMeanControlsSD,PitchMeanSchizo,PitchMeanSchizoSD,PitchSDControls,PitchSDControlsSD,PitchSDSchizo,PitchSDSchizoSD)

DATA=rbind(data11,data)

```

```{r}
#effectsize for own data
EffectSizeMean=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = DATA)

# making lm
ModelMEANDATA = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=EffectSizeMean, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(ModelMEANDATA)

resMDATA = rma(yi, vi, data = EffectSizeMean, slab=Article)
forest(resMDATA)

#counting participants
SubsetMDATA=subset(EffectSizeMean[EffectSizeMean$vi>0,])
sum(SubsetMDATA$SampleSizeSchizo, na.rm=T)
sum(SubsetMDATA$SampleSizeContros, na.rm=T)

#----------------------------

#SD
SDDATA=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = DATA)

# making lm
ModelSDDATA = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=SDDATA, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(ModelSDDATA)

resSDDATA = rma(yi, vi, data = SDDATA, slab=Article)
forest(resSDDATA)

SubsetSDDATA=subset(SDDATA[SDDATA$vi>0,])
sum(SubsetSDDATA$SampleSizeSchizo, na.rm=T)
sum(SubsetSDDATA$SampleSizeContros, na.rm=T)
```
#####Reporting

######Mean
5 studies
241 participants with schizophrenia and 169 comparison participants.
The overall estimated difference (Cohen’s d) in pitch variability between the schizophrenia and the comparison groups was .17, SE= 0.13
######SD
7 studies
332 participants with schizophrenia and 230 comparison participants.
The overall estimated difference (Cohen’s d) in pitch variability between the schizophrenia and the comparison groups was -0.12, SE= 0.43


3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.
```{r}
# do summary of model(the rm2(yi, Vi,data, slab='study')), => gives i^2 and TAU^2(how much var. between study).
# Find TAU and I^2 from Mean
summary(resMDATA)
#TAU^2=0.02 (SE=0.055)
#I^2 26.15%

#funnelplot
funnel(resMDATA, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
regtest(resMDATA)
ranktest(resMDATA)

?ranktest()

#influential...
infM <- influence(resMDATA)
print(infM)
plot(infM)

#Martinez et al 2015 is significant in the analysis of influential studies, which indicates that it is an outlier, and thus this study affects the analysis... funnelplot

#----------------------

# Find TAU and I^2 from SD
summary(resSDDATA)
#TAU^2=3.49 (SE=2.06)
#I^2 98.18%

#TAU^2(estimated amount of total heterogeneity) is = 3.49 (SE=2.06), while I^2 (total heterogeneity / total variability) is = 98.18%. These high values indicate a relatively high amount of spreaded data. Cohen et al's study, is also far away....

#funnel plot
funnel(resSDDATA, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
regtest(resSDDATA) #reg test is significant (P=.003), could indicate a publication bias. 
ranktest(resSDDATA) "rank indicates 0"

?influence()

#influen...
infSD <- influence(resSDDATA)
print(infSD)
plot(infSD)

#Cohen et al. 2014 is significant in the analysis of influential studies, which indicates that it is an outlier, and thus this study affects the analysis... funnelplot


#desto højere i^2 værdi, desto mere spredelse.




```


## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=0
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2


